
\newpage
\selectlanguage{english}

\begin{center}
\Large \textbf{Dedication}
\end{center} 
\vfill
\begin{center}
To my beloved children Karla and Matěj and my dearest lovely wife Marie,

without whom this work would not have been completed.
\end{center} 
\vfill
\newpage
\mbox{}
\newpage
\pagestyle{plain} % other pages with number
\begin{center}
\Large \textbf{Acknowledgements}
\end{center} 
I would like to express my gratefulness to my supervisor Ing. Milan Šárek, CSc. for his support and guidance throughout the beginning of my research. 
I would also like to thank the staff of association CESNET, especially Ing. Jiří Navrátil, CSc., who gave me an opportunity to work with computing resources connected via high speed academic network and to follow the development of the networking technologies. Further thanks go to the staff of METACENTRUM, department of CESNET, for giving support and access to powerful computing infrastructure for scientific purposes. I would like to thank to the staff of EGI.eu, who gave me an opportunity to be EGI Champion, selected users of scientific European grid-computing and cloud-computing infrastructure. This made me possible to meet the most inspiring people in the technology domain as well as scientific domain face to face. 

I would like to thank the fellows from Musical Acoustic Research Center of Academy of Performing Arts, especially RNDr. Marek Frič, PhD. and Ing. Zdeněk Otčenášek, Phd. for their inspiring atmosphere in the research domain of voice science. 

I would like to thank my colleagues from Laboratory of Biocybernetics and Computer Aided Teaching within Institute of Pathological Physiology including Mgr. Marek Mateják, Ing. Jan Šilar, Ing. Filip Ježek, Ing. Martin Tribula and MUDr. et Mgr. Pavol Privitzer, who provided me pleasure and inspiring atmosphere for my research and development work during the second part of my research. Especially, I would like to thank to my consultant Doc. MUDr. Jiří Kofránek, CSc. and a fellow MUDr. Stanislav Matoušek, PhD., who taught me the subject of medical physiology and who guided me within the topic of mathematical models and simulation of integrative physiology. Last but not least, thanks to Veronika Sýkorová, DiS and Klára Ulčová, DiS for their help with graphical schemas. 

The research projects were solved within the cooperation between association CESNET, First Faculty of Medicine of Charles University in Prague and Academy of Performing Arts in Prague and was partially supported by the project of Ministry of Education, Youth and Sport of the Czech Republic MSM6383917201, by the projects of Development Fund of CESNET 2009/361, 2010/384, 2011/423, 2011/431 and by the project of Ministry of Industry and Trade of the Czech Republic MPO FR-TI3/869.

\newpage
\mbox{}
